---
bibliography: [references.bib]
nocite: '@*'
csl: 'apa.csl'
output:
  pdf_document:
    includes:
      before_body: title-page.sty
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
\textbf{\LARGE Abstract}

\newpage

\newpage
\textbf{\LARGE Introduction}

In the past decade, the usage of public rental bikes has increased in various 
urban cities. In cities, rental bikes have become a crucial part of transportation
because of many reasons. They provide free or affordable access to transportation
for short-distance trips instead of requiring a private vehicle and help reduce 
congestion, noise and air pollution in a city (Winters, 2020). Bike Sharing systems
are now rapidly growing across the world at nearly 2000 various operating programs
in total as of May 2021 (Yu et al., 2021). The demand for these systems is growing
at 14.3% compounded annually from 2017 to 2025 (Cooper, 2019). 
 
There are various public bike rental systems across the world, like in the city 
of Toronto where they have approximately 6,850 bikes and its users have completed
2.9 million trips as of 2020 (Bike Share Toronto, 2021). In New York City there 
are "... nearly eight hundred thousand (773,000) ride a bicycle regularly" and 
have completed over 7 million rides as of September 2021 with a 26% growth in 
daily cycling between 2014 to 2019 (New York City DOT, 2021). One of the biggest
bike-sharing systems is found in China at Hangzhou and Wuhan where their residents
have access to 90,000 and 70,000 bicycles respectively 
(Borowska-Stefańska et al., 2021). 
 
Like many other cities within the world, the city of Seoul in South Korea has 
also founded a bike-sharing system called _Ttareungyi_ or Seoul Public Bike 
started in 2015. It was started to " resolve issues of traffic congestion, air 
pollution, and high oil prices in Seoul, and to build a healthier society while 
enhancing the quality of life for Seoul citizens" (Seoul Metropolitan Government,
2015). As of March 20, 2018, it has exceeded 620,000 memberships with 38% of the 
users using the bikes during rush hour (Seoul Metropolitan Government, 2018). Due
to the high usage of bikes in cities effective management of bike-sharing systems 
must be developed to ensure the general public can access the service when required.
 
The following data set from the Seoul Public Bike initiative provides various 
information like the number of bikes rented, temperature, humidity, wind speed, 
visibility, solar radiation, snowfall and more. This can be used to see how 
weather affects the number of bikes rented per hour and makes a predictive model
for future uses from the given data from December 1st, 2017 to December 1st, 2018. 
This is a very important question to consider since an online survey conducted 
within the geographical region of Asia in 2016 identified,  climate as an "... 
important physical barrier with 25% of respondents strongly agreeing that it 
constrains them physically, particularly warm or warmth and high precipitation 
levels."(Mateo-Babiano et al., 2017) among other reasons like infrastructure 
issues and others.

\newpage


```{r echo=FALSE}

library("readxl")
myData = read.csv(file='SeoulBikeData.csv', check.names = F)

colnames(myData) = c("Date", "RentedBikeCount", "Hour", "Temperature", "Humidity", "Windspeed", "Visibility", "DewPointTemp", "SolarRadiation", "Rainfall", "Snowfall", "Seasons", "Holiday", "FunctioningDay")
n = dim(myData)[1]
X0 = as.Date(myData$Date, format = "%d/%m/%Y")
  
Y = myData$RentedBikeCount
X1 = myData$Hour
X2 = myData$Temperature
X3 = myData$Humidity
X4 = myData$Windspeed
X5 = myData$Visibility
X6 = myData$DewPointTemp
X7 = myData$SolarRadiation
X8 = myData$Rainfall
X9 = myData$Snowfall
#X10 (seasons)comprises of 3 dummy variables, as done in lec 15.
X10 = myData$Seasons
X11 = myData$Holiday
X12 = myData$FunctioningDay


```

```{r echo=FALSE}
DayOfTheWeek = format(X0, "%a")
weekends = as.numeric(as.numeric(DayOfTheWeek == "Sat") != as.numeric(DayOfTheWeek == "Sun"))
# this assigns = 1 when DayOfTheWeek == "Sat" or "Sun"
```

```{r echo=FALSE}

fit = lm (Y ~ X1 + X2 + X3 + X4 + X5 + X7 + X8 + X9)

fit2 = lm (Y ~ X1 + X2 + X3 + X5 + X7 + X8 + X9)

library(leaps)

regsubset = regsubsets(Y ~ X1 + X2 + X3 + X4 + X5 + X7 + X8 + X9, nbest=4, data = myData)
aprout = summary(regsubset)
n = length(Y)
p = apply(aprout$which, 1, sum)
aprout$aic = aprout$bic - log(n) * p + 2 * p
with(aprout, round(cbind(which, rsq, adjr2, cp, bic, aic), 3))

```

```{r echo=FALSE}
model2 = lm(Y ~ X1 + X2 + X3 + X5 + X7 + X8 + X9)
anova(fit2, model2)

```
 Remove windspeed (X4) due to p-value > alpha when we do anova()


Now, to test for normality, we will check whether the normal Q-Q plot (normal probability plot) of the residual plot is close to the line y=x or not. 


```{r echo=FALSE}

qqnorm(fit$residuals)
qqline(fit$residuals)

```

We see that the normal Q-Q plot deviates from the line y=x on either sides. Therefore, normality does not hold. Therefore, we will conduct a box-cox transformation to make sure the model 

```{r echo=FALSE}
library(MASS)

myData_mod = subset(myData, myData$RentedBikeCount > 0) # deleting rows with Y <= 0 as box cox requires all response Yi's to be > 0

Y = myData_mod$RentedBikeCount
X1 = myData_mod$Hour
X2 = myData_mod$Temperature
X3 = myData_mod$Humidity
X4 = myData_mod$Windspeed
X5 = myData_mod$Visibility
X6 = myData_mod$DewPointTemp
X7 = myData_mod$SolarRadiation
X8 = myData_mod$Rainfall
X9 = myData_mod$Snowfall
#X10 (seasons)comprises of 3 dummy variables, as done in lec 15.
X10 = myData_mod$Seasons
X11 = myData_mod$Holiday
X12 = myData_mod$FunctioningDay

X0 = as.Date(myData_mod$Date, format = "%d/%m/%Y")
DayOfTheWeek = format(X0, "%a")
weekends = as.numeric(as.numeric(DayOfTheWeek == "Sat") != as.numeric(DayOfTheWeek == "Sun"))
# this assigns = 1 when DayOfTheWeek == "Sat" or "Sun"

fit2 = lm (Y ~ X1 + X2 + X3 + X4 + X5 + X7 + X8 + X9)

result = boxcox(fit2)

lambda = result$x[which.max(result$y)]

K2 = prod(Y^(1/length(Y))) # prod(y)^(1/length(y) = prod(y^(1/length(y))) -- check formula

K1 = 1 / (lambda*K2^(lambda-1))

yt = (((Y^lambda)-1)/K1)

new_model = lm(yt ~ X1 + X2 + X3 + X4 + X5 + X7 + X8 + X9)


qqnorm(new_model$residuals)
qqline(new_model$residuals)


```

```{r echo=FALSE}


plot(fit$fitted.values, fit$residuals)

plot(new_model$fitted.values, new_model$residuals)

hist(fit$residuals)
hist(new_model$residuals)

```


Also,

As there are 4 seasons, we will have 3 dummy variables.

Let us take the Base category for X10: Autumn

D1 = {
1   , if season = Winter
0   , o/w
}

D2 = {
1   , if season = Spring
0   , o/w
}

D3 = {
1   , if season = Summer
0   , o/w
}

```{r echo=FALSE}

D1 = as.numeric(X10 == "Winter")
D2 = as.numeric(X10 == "Spring")
D3 = as.numeric(X10 == "Summer")

model_with_season = lm(yt ~ X1 + X2 + X3 + X4 + X5 + X7 + X8 + X9 + D1 + D2 + D3
                       + D1:X1 + D1:X2 + D1:X3 + D1:X4 + D1:X5 + D1:X7 + D1:X8 + D1:X9
                       + D2:X1 + D2:X2 + D2:X3 + D2:X4 + D2:X5 + D2:X7 + D2:X8 + D2:X9
                       + D3:X1 + D3:X2 + D3:X3 + D3:X4 + D3:X5 + D3:X7 + D3:X8 + D3:X9)

F_test = (0.059039/25)/(0.005287)^2

qf(0.05, 25, 8431)

```

The seasons should be kept in the model, as p-value << alpha. REJECT H_0 that states that seasons do not depend on the response variable Y.


Now, we try to fit a model which includes the weekend variable as well as the seasons

```{r echo=FALSE}


model_with_weekend_var = lm(yt ~ X1 + X2 + X3 + X5 + X7 + X8 + X9
                            + weekends:X1 + weekends:X2 + weekends:X3
                            + weekends:X5 + weekends:X7
                            + weekends:X8 + weekends:X9
                            + D1 + D2 + D3 + weekends
                            + D1:X1 + D1:X2 + D1:X3 + D1:X5 + D1:X7 + D1:X8 + D1:X9
                            + D2:X1 + D2:X2 + D2:X3 + D2:X5 + D2:X7 + D2:X8 + D2:X9
                            + D3:X1 + D3:X2 + D3:X3 + D3:X5 + D3:X7 + D3:X8 + D3:X9)

james_model = lm(formula = yt ~ (X1 + X2 + X3 + X5 + X7 + X8 + X9) * (D1 + D2 + D3 + weekends))

months = format(X0, "%b")

m1 = as.numeric(months == "Jan")
m2 = as.numeric(months == "Feb")
m3 = as.numeric(months == "Mar")
m4 = as.numeric(months == "Apr")
m5 = as.numeric(months == "May")
m6 = as.numeric(months == "Jun")
m7 = as.numeric(months == "July")
m8 = as.numeric(months == "Aug")
m9 = as.numeric(months == "Sep")
m10 = as.numeric(months == "Oct")
m11 = as.numeric(months == "Nov")


# LET BASE CATEGORY = December

months_transformation_model = lm(formula = yt ~ (X1 + X2 + X3 + X5 + X7 + X8 + X9) * (weekends 
                                                                                    + m1 + m2 + m3 + m4 + m5 + m6
                                                                                    + m7 + m8 + m9 + m10 + m11)
                               )
```

```{r echo=FALSE}

h0 = as.numeric(X1 == 0)
h1 = as.numeric(X1 == 1)
h2 = as.numeric(X1 == 2)
h3 = as.numeric(X1 == 3)
h4 = as.numeric(X1 == 4)
h5 = as.numeric(X1 == 5)
h6 = as.numeric(X1 == 6)
h7 = as.numeric(X1 == 7)
h8 = as.numeric(X1 == 8)
h9 = as.numeric(X1 == 9)
h10 = as.numeric(X1 == 10)
h11 = as.numeric(X1 == 11)
h12 = as.numeric(X1 == 12)
h13 = as.numeric(X1 == 13)
h14 = as.numeric(X1 == 14)
h15 = as.numeric(X1 == 15)
h16 = as.numeric(X1 == 16)
h17 = as.numeric(X1 == 17)
h18 = as.numeric(X1 == 18)
h19 = as.numeric(X1 == 19)
h20 = as.numeric(X1 == 20)
h21 = as.numeric(X1 == 21)
h22 = as.numeric(X1 == 22)


# LET BASE CATEGORY : X1 = 23

hours_transformation_model = lm(formula = yt ~ (X2 + X3 + X5 + X7 + X8 + X9) * (weekends 
                                                                                    + h0 + h1 + h2 + h3 + h4 + h5 
                                                                                    + h6 + h7 + h8 + h9 + h10 + h11
                                                                                    + h12 + h13 + h14 + h15 + h16 + h17
                                                                                    + h18 + h19 + h20 + h21 + h22
                                                                                    )
                               )

mon_hr_transformation_model = lm(formula = yt ~ (X2 + X3 + X5 + X7 + X8 + X9) * (weekends 
                                                                                    + h0 + h1 + h2 + h3 + h4 + h5 + h6
                                                                                    + h7 + h8 + h9 + h10 + h11
                                                                                    + h12 + h13 + h14 + h15 + h16 + h17
                                                                                    + h18 + h19 + h20 + h21 + h22
                                                                                    + m1 + m2 + m3 + m4 + m5 + m6
                                                                                    + m7 + m8 + m9 + m10 + m11
                                                                                    )
                               )


#BEST MODEL UNTIL NOW BELOW:
season_hr_model = lm(formula = yt ~ (X2 + X3 + X5 + X7 + X8 + X9) * (weekends 
                                                                                    + h0 + h1 + h2 + h3 + h4 + h5 + h6
                                                                                    + h7 + h8 + h9 + h10 + h11
                                                                                    + h12 + h13 + h14 + h15 + h16 + h17
                                                                                    + h18 + h19 + h20 + h21 + h22
                                                                                    + D1 + D2 + D3
                                                                                    ))

#as the R^2 adj for below has a very low value, due to the drop in it we see that HOURS should be included in the model.
month_season_model = lm(formula = yt ~ (X2 + X3 + X5 + X7 + X8 + X9) * (weekends    + m1 + m2 + m3 + m4 + m5 + m6
                                                                                    + m7 + m8 + m9 + m10 + m11
                                                                                    + D1 + D2 + D3
                                                                                    ))
```


```{r echo=FALSE}

X11 = myData_mod$Holiday

isHoliday = as.numeric(X11 == "Holiday")

season_hr_holiday_model = lm(formula = yt ~ (X2 + X3 + X5 + X7 + X8 + X9) * (weekends + isHoliday
                                                                                    + h0 + h1 + h2 + h3 + h4 + h5 + h6
                                                                                    + h7 + h8 + h9 + h10 + h11
                                                                                    + h12 + h13 + h14 + h15 + h16 + h17
                                                                                    + h18 + h19 + h20 + h21 + h22
                                                                                    + D1 + D2 + D3
                                                                                    ))

```

There is not a significant increase in the adjusted R-squared value when we include the Holiday dummy variable. Therefore, our final model is season_hr_model, Y = DO IN LATEX

\newpage
\textbf{\LARGE Bibliography}